---
title: 'ALY6010: R Practice 2'
subtitle: 'Module 2'
author: "Shamim Sherafati" 
date: "2022-11-09"
output:
  html_document: 
    toc: yes
    keep_md: yes
  pdf_document: default
---

### Importing packages
```{r}
#install.packages("plyr")
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("ggpubr")
#install.packages("ggplot2")
#install.packages("plotly")
#install.packages("moments")
#install.packages('gmodels')
#install.packages("stringr")

library(plyr) 
library(dplyr)
library(tidyr)
library(tidyverse)
library(psych)
library(ggpubr)
library(ggplot2)
library(plotly)
library(moments)
library(gmodels)
library(stringr)
```
### Import the file *Student Survey.csv*

```{r}
StudentMH = read.csv("~/Documents/ALY 6010/dataset/Student Mental health.csv")
StudentMH
```
### Dimension of the dataset
```{r}
dim(StudentMH) #get dim to gain information before start cleaning data
```

### Data Cleaning

```{r}
# Sorted dataframe with descending age
StudentMH_sort <- StudentMH[order(-StudentMH$Age),]
head(StudentMH_sort)


# Rename column 
names(StudentMH_sort)[names(StudentMH_sort) == 'Choose.your.gender'] <- 'Gender'
names(StudentMH_sort)[names(StudentMH_sort) == 'What.is.your.course.'] <- 'Course'
names(StudentMH_sort)[names(StudentMH_sort) == 'Your.current.year.of.Study'] <- 'Current.study.year'
names(StudentMH_sort)[names(StudentMH_sort) == 'What.is.your.CGPA.'] <- 'CGPA'
names(StudentMH_sort)[names(StudentMH_sort) == 'Do.you.have.Depression.'] <- 'Depression'
names(StudentMH_sort)[names(StudentMH_sort) == 'Do.you.have.Anxiety.'] <- 'Anxiety'
names(StudentMH_sort)[names(StudentMH_sort) == 'Do.you.have.Panic.attack.'] <- 'Panic.attack'
names(StudentMH_sort)[names(StudentMH_sort) == 'Did.you.seek.any.specialist.for.a.treatment.'] <- 'Specialist.Treatment'
head(StudentMH_sort)

# Drop column "Timestamp"
StudentMH_drop <- select(StudentMH_sort, -1)
head(StudentMH_drop)

```


```{r}
str(StudentMH_drop)
```

```{r}
summary(StudentMH_drop) #get summary of dataset after cleaning data
```

### Mean, Min, Max, STD 

```{r}
#get the Mean, Min, Max, STD of each current year of study/course
StudentMH_drop %>%
group_by(Current.study.year)%>%
summarise_at(vars(Course),list(Mean=mean, Min=min, Max=max,STD=sd))

#get the Mean, Min, Max, STD of each course/ CGPA
StudentMH_drop %>%
group_by(Course)%>%
summarise_at(vars(CGPA),list(Mean=mean, Min=min, Max=max,STD=sd))

```

### Subsets of data

```{r}
#subset of data for finding the Anxiety, Depression and need of Specialist.Treatment
subset(StudentMH_drop, Anxiety == "Yes") #Find subset of having Anxiety 
subset(StudentMH_drop, Specialist.Treatment == "Yes") #subset for need of Specialist.Treatment
subset(StudentMH_drop, Depression == "Yes") #subset of having Depression 

#subset of data for finding the courses and GPA for first year of study
subset(StudentMH_drop, Current.study.year == "year 1") 

```
### Describe Data

```{r}
psych::describe(StudentMH_drop, na.rm = TRUE, interp=FALSE,skew = TRUE, ranges = TRUE,trim=.1,
type=3,check=TRUE,fast=NULL,quant=NULL,IQR=FALSE) #describe dataset 

describeData(StudentMH_drop,head=1,tail=2) #describe dataset for first head and two of tails
```
### Table For Course and Current.study.year

```{r}
#table for different courses which must taken each years
table(StudentMH_drop$Course, StudentMH_drop$Current.study.year)
```
### Frequency Tables For age and  Marital status

```{r}
#Frequency Table for marital status and age
StudentMH_Ftable <- table(StudentMH_drop$Age, StudentMH_drop$Marital.status) 
ftable(StudentMH_Ftable)
```
### Cross Table for Specialist.Treatment, Depression, Panic.attack

```{r}
#create CrossTable for 'Depression' and 'Panic.attack'
CrossTable(StudentMH_drop$Depression, StudentMH_drop$Panic.attack,
                 dnn = c("Depression", "Panic.attack"))
```
```{r}
#create CrossTable for 'Depression' and 'Specialist.Treatment'
CrossTable(StudentMH_drop$Depression, StudentMH_drop$Specialist.Treatment,
                 dnn = c("Depression", "Specialist.Treatment"))
```
```{r}
#create CrossTable for 'Specialist.Treatment' and 'Panic.attack'
CrossTable(StudentMH_drop$Specialist.Treatment, StudentMH_drop$Panic.attack,
                 dnn = c("Specialist.Treatment", "Panic.attack"))
```
### Three-line table

```{r}
#install.packages("bruceR")
library(bruceR)

print_table(
  StudentMH_drop,
  nspaces = 1,
  row.names = TRUE,
  col.names = TRUE,
  title = "",
  note = "",
  append = "",
  line = TRUE,
  file = NULL,
  file.align.head = "auto",
  file.align.text = "auto")   #create three line table for dataset
```

### R abline() and par() function

```{r}

StudentMH_zone <- StudentMH_drop$Age #create new variable
par(mfrow=c(2,2))  #use par function to create histogram and boxplot

hist(StudentMH_zone) #histogram for new variable
boxplot(StudentMH_zone, horizontal=TRUE) #boxplot for new variable with par() function

set.seed(1234); StudentMH_drop<-rnorm(200)
hist(StudentMH_drop, col="lightblue")
abline(v = mean(StudentMH_drop), col="red", lwd=3, lty=2) #use abline function for histogram
```
### Different courses GPA Score

```{r}
#ggplot for education and marital status

ggplot (data = df$StudentMH_drop) + geom_bar(mapping = aes(x=CGPA, fill=Course), position="dodge") + labs(title="Course GPA", x ="Course", y = "GPA")
```
### Average of ages who go for Specialist.Treatment

```{r}
#Average of ages who go for Specialist.Treatment
TopAge <- StudentMH_drop

TreatmentAge <- aggregate(Age ~ Specialist.Treatment, TopAge, mean)
slice( TreatmentAge[order(-TreatmentAge$Age),], 1:5)
```
```{r}
#create box plot for Average of ages who go for Specialist.Treatment based on the previous code

boxplot( TopAge$Age ~ TopAge$Specialist.Treatment, main="Average of ages who go for Specialist.Treatment", xlab = "Specialist.Treatment", ylab = "Age ", col = "light pink")
```
### Compare different ages in different year of study

```{r}
#create box plot for different ages in different year of study

box1 <- boxplot( StudentMH_drop$Age ~ StudentMH_drop$Current.study.year, main="Compare different ages in different year of study", xlab = "Current.study.year", ylab = "Age ", col = "light yellow")
```
### Proporation for each course based on Genders

```{r}
#ggplot for Gender based on courses
ggplot(data = StudentMH_drop) + geom_bar(mapping = aes(x=Gender, fill=Course), position="dodge") + labs(title="Different courses/Gender", x ="Course", y = "Gender")
```
### Relation between Depression and Anxiety

```{r}
#creat jitter chart for Depression and Anxiety
# Use single color
ggplot(StudentMH_drop, aes(x=Depression, y=Anxiety)) + 
    geom_jitter(position=position_jitter(0.2), color="red") + geom_boxplot(alpha = 0)
# Change stripchart colors by groups
p<-ggplot(StudentMH_drop, aes(x=Depression, y=Anxiety, color=dose)) +
  geom_jitter(position=position_jitter(0.2))

```
### Proportion of those habe panik attack attend to Specialist.Treatment 

```{r}
# Change stripchart colors by groups
plot1 <- ggplot(StudentMH_drop, aes(x=Panic.attack, y=Specialist.Treatment, color="Panic.attack")) +
  geom_jitter(position=position_jitter(0.2)) + geom_boxplot(alpha = 0)
# Change the position : interval between stripchart of the same group
p<-ggplot(StudentMH_drop, aes(x=Panic.attack, y=Specialist.Treatment, color=supp, shape=supp)) +
  geom_jitter(position=position_dodge(0.10))


# Change stripchart colors by groups
plot2 <- ggplot(StudentMH_drop, aes(x=Anxiety, y=Specialist.Treatment, color="Anxiety")) +
  geom_jitter(position=position_jitter(0.2)) + geom_boxplot(alpha = 0)
# Change the position : interval between stripchart of the same group
p<-ggplot(StudentMH_drop, aes(x=Anxiety, y=Specialist.Treatment, color=supp, shape=supp)) +
  geom_jitter(position=position_dodge(0.10))


plot_grid(plot1, plot2, labels = "AUTO")   #combine two charts together
```

### Different ages gender attend to Specialist.Treatment

```{r}
#install.packages("scatterPlotMatrix")
#library(scatterPlotMatrix)

#create ggplot for age and contact-info
ggplot(data = StudentMH_drop) + geom_bar(mapping = aes(x=Specialist.Treatment, fill= Gender), position="dodge") + labs(title="Different ages attend to Specialist.Treatment", x ="Specialist.Treatment", y = "Gender")

```
### Impact of anxiety on different ages

```{r}
#install.packages("cowplot")
library(cowplot)

StudentMH_drop1 <- ggplot(StudentMH_drop, aes(x = Age, y = Anxiety)) +
  geom_boxplot() + theme_bw()    #boxplot for anxiety on different ages

StudentMH_drop2 <- ggplot(StudentMH_drop, aes(x = Anxiety, fill = Age)) +
  geom_density(alpha = 0.7) + theme_bw() +   #create density
  theme(legend.position = c(0.8, 0.8))   

plot_grid(StudentMH_drop1, StudentMH_drop2, labels = "AUTO")   #combine two charts together
```
#### Compare Ages and gender based on their current study year

```{r}
#install.packages("patchwork")
library(ggplot2)
library(patchwork)

p1 <- ggplot(StudentMH_drop) + geom_point(aes(Gender, Current.study.year)) #create chart for year and gender
p2 <- ggplot(StudentMH_drop) + geom_boxplot(aes(Age, Current.study.year, group = Current.study.year)) #create chart for year and ages

plot_grid(p1, p2, labels = "AUTO")   #combine two charts together

```
### check if those with depression, anxiety attend to Specialist.Treatment or not

```{r}
q1 <- ggplot(StudentMH_drop) + geom_point(aes(Depression, Specialist.Treatment))
q2 <- ggplot(StudentMH_drop) + geom_boxplot(aes(Anxiety, Specialist.Treatment, group = Specialist.Treatment))
#q3 <- ggplot(StudentMH_drop) + geom_smooth(aes(Panic.attack, Specialist.Treatment, group = Panic.attack))
q4 <- ggplot(StudentMH_drop) + geom_bar(aes(Specialist.Treatment))

#install.packages("multipanelfigure")
library(magrittr)
library(multipanelfigure)
figure1 <- multi_panel_figure(columns = 2, rows = 2, panel_label_type = "none")

figure1 %<>%
  fill_panel(q1, column = 1, row = 1) %<>%
  fill_panel(q2, column = 2, row = 1) %<>%
  #fill_panel(q3, column = 1, row = 2) %<>%
  fill_panel(q4, column = 1, row = 2)
figure1
```
### References
Aphalo, Pedro J. 2017. Ggpmisc: Miscellaneous Extensions to ’Ggplot2’. https://CRAN.R-project.org/package=ggpmisc.

Attali, Dean. 2017. GgExtra: Add Marginal Histograms to ’Ggplot2’, and More ’Ggplot2’ Enhancements. https://github.com/daattali/ggExtra.

Wilke, Claus O. 2017. Ggridges: Ridgeline Plots in ’Ggplot2’. https://CRAN.R- project.org/package=ggridges.

Wickham, Hadley, and Winston Chang. 2017. Ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics.

Schloerke, Barret, Jason Crowley, Di Cook, Francois Briatte, Moritz Marbach, Edwin Thoen, Amos Elberg, and Joseph Larmarange. 2016. GGally: Extension to ’Ggplot2’. https://CRAN.R-project.org/package=GGally.



### Appendix
```{r code=readLines(knitr::purl('~/Documents/ALY 6010/Week 2/Module 2- R practice.Rmd', documentation = 0)), eval = FALSE}
```






