---
title: 'ALY6010: R Practice 3'
subtitle: 'Module 3'
author: "Shamim Sherafati" 
date: "2022-11-20"
output:
  html_document: 
    toc: yes
    keep_md: yes
  pdf_document: default
---

### Importing packages
```{r}
#install.packages("plyr")
#install.packages("dplyr")
#install.packages("tidyr")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("ggpubr")
#install.packages("ggplot2")
#install.packages("plotly")
#install.packages("moments")
#install.packages('gmodels')
#install.packages("stringr")


library(plyr) 
library(dplyr)
library(tidyr)
library(tidyverse)
library(psych)
library(ggpubr)
library(ggplot2)
library(plotly)
library(moments)
library(gmodels)
library(stringr)
```

### Importing data
```{r}
FoodSurvey = read.csv("~/Documents/Northeastern University/ALY 6010/dataset/Food Survey - Data.csv", header = TRUE)
FoodSurvey
```
### Initial Summary of Data set
```{r}
summary(FoodSurvey) #Initial summary
```
### Data Frame
```{r}
df_foodsurvey <- as.data.frame(read.csv("~/Documents/Northeastern University/ALY 6010/dataset/Food Survey - Data.csv", header = TRUE))
df_foodsurvey
```

### Data Cleaning
```{r}
# Rename column 
names(FoodSurvey)[names(FoodSurvey) == 'DHBreakfast'] <- 'breakfast.in.DH'
names(FoodSurvey)[names(FoodSurvey) == 'NDHBreakfast'] <- 'breakfast.NOTin.DH'
names(FoodSurvey)[names(FoodSurvey) == 'BClass'] <- 'breakfast.in.Class'
names(FoodSurvey)[names(FoodSurvey) == 'DHBoxes'] <- 'BoxOfFood.in.DH'
names(FoodSurvey)[names(FoodSurvey) == 'NDHBoxes'] <- 'BoxOfFood.NOTin.DH'
names(FoodSurvey)[names(FoodSurvey) == 'NDHMeals'] <- 'Meal.in.Dorm'

head(FoodSurvey)

# Drop column "Timestamp "
FoodSurvey_drop <- select(FoodSurvey, -1)
head(FoodSurvey_drop)

# delete 'NA' value
Data_FoodSurvey <- na.omit(FoodSurvey_drop)

head(Data_FoodSurvey)
```
### Final Summary
```{r}
summary(Data_FoodSurvey) #get final summary after cleaning data
```


### string of data after cleaning
```{r}
str(Data_FoodSurvey)
```
### Summary of Nutrition data

```{r}
summary(Data_FoodSurvey$Nutrition)
```

### A. One Sample t-test of Nutrition with two sided tailed

```{r}
mean(Data_FoodSurvey$Nutrition) #get the mean of Nutrition
sd (Data_FoodSurvey$Nutrition) #get the Standard Deviation of Nutrition

t.test(Data_FoodSurvey$Nutrition, mu = 0, alternative = "two.sided")

#If the p-value in a null hypothesis is greater than the significance level, then the null hypothesis is accepted
#If the p-value in an alternate hypothesis is smaller than the significance level, then the alternative hypothesis is accepted
```
### Visualize Nutrition using box plots for test-A

```{r}
ggboxplot(Data_FoodSurvey$Nutrition, 
          ylab = "Nutrition", xlab = FALSE,
          ggtheme = theme_minimal())
```

### Preliminary  test to check one-sample t-test assumptions for test-A

```{r}
shapiro.test(Data_FoodSurvey$Nutrition) # => p-value < 2.2e-16
```

### Visual inspection for test-A

```{r}
ggqqplot(Data_FoodSurvey$Nutrition, ylab = "Nutrition",
         ggtheme = theme_minimal()) #From the normality plots, we conclude that the data may come from normal distributions.
```

### Compute one-sample t-test for test-A

```{r}
#We want to know, if the average Nutrition of the students differs from 5 (two-tailed test)?
# One-sample t-test
test1 <- t.test(Data_FoodSurvey$Nutrition, mu = 5)
# Printing the results
test1 
```

### B. One Sample t-test of Having Meal.in.Dorm with right tailed

```{r}
mean(Data_FoodSurvey$Meal.in.Dorm) #get the mean of Meal.in.Dorm
sd (Data_FoodSurvey$Meal.in.Dorm) #get the Standard Deviation of Meal.in.Dorm

#We want to know, if the average Nutrition of the students greater than 10 (greater-tailed test)?
test2 <- t.test(Data_FoodSurvey$Meal.in.Dorm, mu = 10, alternative = 'greater') #One Sample t-test of Meal.in.Dorm
test2
```

### Visualize of Having Meal.in.Dorm using box plots for test-B

```{r}
ggboxplot(Data_FoodSurvey$Meal.in.Dorm, 
          ylab = "Meal.in.Dorm", xlab = FALSE,
          ggtheme = theme_minimal())
```

### Preliminary  test to check one-sample t-test assumptions for test-B

```{r}
shapiro.test(Data_FoodSurvey$Meal.in.Dorm) # => p-value = 1
```

### Visual inspection for test-B

```{r}
ggqqplot(Data_FoodSurvey$Meal.in.Dorm, ylab = "Meal in Dorm",
         ggtheme = theme_minimal()) #From the normality plots, we conclude that the data may come from normal distributions.
```

### C. One Sample t-test of  breakfast.in.Dinng Hall with left tailed

```{r}
mean(Data_FoodSurvey$breakfast.in.DH) #get the mean of breakfast.in.Dinng Hall
sd(Data_FoodSurvey$breakfast.in.DH) #get the standard deviation of breakfast.in.Dinng Hall

##We want to know, if the average Nutrition of the students less than 7 (less-tailed test)?
test3 <- t.test(Data_FoodSurvey$breakfast.in.DH, mu = 7, alternative = 'less') #One Sample t-test of breakfast.in.DH with 
test3
```

### Visualize of having breakfast in dining hall using box plots for test-C

```{r}
ggboxplot(Data_FoodSurvey$breakfast.in.DH, 
          ylab = "breakfast in dining hall", xlab = FALSE,
          ggtheme = theme_minimal())
```

### Preliminary  test to check one-sample t-test assumptions for test-C

```{r}
shapiro.test(Data_FoodSurvey$breakfast.in.DH) # => p-value < 2.2e-16
```

### Compute one-sample t-test for test-C

```{r}
#We want to know, if the average of breakfast.in.DH of the students less than 7 (left-tailed test)?
# One-sample t-test
test2 <- t.test(Data_FoodSurvey$breakfast.in.DH, mu = 7)
# Printing the results
test2
```

### Hypothesis testing for p-value

```{r}
A <- Data_FoodSurvey$BoxOfFood.NOTin.DH  #create data A for one variable n dataset

B <- Data_FoodSurvey$breakfast.NOTin.DH # #create data B for another variable n dataset

wilcox.test(A,B, paired = TRUE, correct = TRUE)  #Wilcoxon rank sum exact test for A and B 

```

### P-value test for Nutrition

```{r}
t = (mean(Data_FoodSurvey$Nutrition)-2.5)/(sd(Data_FoodSurvey$Nutrition)/sqrt(length(Data_FoodSurvey$Nutrition)))
2*pt(-abs(t),df=length(Data_FoodSurvey$Nutrition)-1)
pt(t, df=136, lower.tail=T)
pt(t, df=136, lower.tail=F)
```

### P-value test for breakfast.in.DH

```{r}
t = (mean(Data_FoodSurvey$breakfast.in.DH)-2.5)/(sd(Data_FoodSurvey$breakfast.in.DH)/sqrt(length(Data_FoodSurvey$breakfast.in.DH)))
2*pt(-abs(t),df=length(Data_FoodSurvey$breakfast.in.DH)-1)
pt(t, df=136, lower.tail=T)
pt(t, df=136, lower.tail=F)
```

### P-value test for Meal.in.Dorm

```{r}
t = (mean(Data_FoodSurvey$Meal.in.Dorm)-2.5)/(sd(Data_FoodSurvey$Meal.in.Dorm)/sqrt(length(Data_FoodSurvey$Meal.in.Dorm)))
2*pt(-abs(t),df=length(Data_FoodSurvey$Meal.in.Dorm)-1)
pt(t, df=136, lower.tail=T)
pt(t, df=136, lower.tail=F)
```

### References

I.	D. Iacobucci, On p-values, Journal of Consumer Research 32 (June 2005), no. 1, 6–12. http://www.journals.uchicago.edu/cgi-bin/resolve?JCR320199PDF 

II.	R. Mason, D. Lind, and W. Marchal, Statistics, An Introduction; 5th edition; Duxbury Press; Brooks/Cole Publishing Company; 1998. 

III.	R. Savage and K. W. Deutsch, A Statistical Model of the Gross Analysis of Transaction Flows Econometrica 28 (1960), no. 3 551–572. 

IV.	Y. M. M. Bishop and S. E. Fienberg, Incomplete Two-Dimensional Contingency Tables, Biometrics 25 (1969), no. 1, 119–128. 




### Appendix
```{r code=readLines(knitr::purl('~/Documents/Northeastern University/ALY 6010/Week 3/Sherafati-Module3-R practice3.Rmd', documentation = 0)), eval = FALSE}
```

