---
title: "Executive Summary Report 2"
output: html_notebook

---

# Shamim Sherafati
### 2022-10-05
## Module 2 -project


1. Print your name 

```{r}
name <- ("Shamim Sherafati")
r_name = paste("Plotting Basics : ",name)
print(r_name)
```
2. Import libraries

```{r}
install.packages ("plyr")
install.packages ("FSA")
install.packages ("FSAdata")
install.packages ("magrittr")
install.packages ("dplyr")
install.packages ("plotrix")
install.packages ("ggplot2")
install.packages ("moments")
```
3. Load the BullTroutRML2 dataset

```{r}
data1 <- read.table (file="~/desktop/BullTroutRML2.txt", header= TRUE, sep=",", stringsAsFactors = FALSE)
data1
```
Describe number 3: Dataset of BullTroutRML2 which was automatically loaded while installing packages.

4. Print the first and last 3 records from the dataset

```{r}
head (data1,  n=3 )
tail (data1, n=3 )
#headtail (data1, n=3)
```
Describe number 4: Used (head) for displaying the first records and used (tail) for displaying the last records with using "n" for the numbers of records which are shown.

5. Filter out all records except those from Harrison Lake

```{r}
library(dplyr)

data_f = filter(data1, lake == 'Harrison')

data_f
```
Describe number 5: For  filtering all datas except one of them, first I introduced an object "data_f" then write "filter" and the name of dataset and used == which means showing only that data which comes after == .

6. Display the first and last 3 records from the filtered dataset

```{r}

head(data_f, n=3)
tail(data_f, n=3)
```
Describe number 6:Its like the previous question with this difference that we should show the head and tail for filtered data, so the object whichi choosed for filtered data, i write.

7. Display the structure of the filtered dataset

```{r}
str(data_f)
```
Describe number 7:for displaying the structure, we should use (str) and then the object which choosed for filtered data which is data_f.

8.Display the summary of the filtered dataset and save it as <t>

```{r}
t <- summary(data_f)
t
```
Describe number 8:First I choosed a name for the summary (t) and then use summary + the name of filtered data.

9. Create a scatterplot for “age” (y variable) and “fl” (x variable) with the following
specifications: 
• Limit of x axis is (0,500)
• Limit of y axis is (0,15)
• Title of graph is “Plot 1: Harrison Lake Trout
• Y axis label is “Age (yrs)”
• X axis label is “Fork Length (mm)

```{r}

plot (data_f$fl, data_f$age, main = ("Plot 1: Horrison Lake Trout"), xlim = c(0, 500), ylim = c(0, 15), xlab = ("Fork Length (mm)"), ylab = ("Age (yrs)"),col="red", cex=0.5, pch=16)

```
Describe number 9:I create a scatter plot with “age” (y variable) and “fl” (x variable). This Scatter Plot shows strong Positive and Linear Relationship between Fork Length and Age of the Dataset.

10. Plot an “Age” histogram with the following specifications:
• Y axis label is “Frequency”
• X axis label is “Age (yrs)”
• Title of the histogram is “Plot 2: Harrison Fish Age Distribution”
• The color of the frequency plots is “cadetblue”
• The color of the Title is “cadetblue”

```{r}

hist(data_f$age, main= "Plot 2: Harrison Fish Age Distribution" , xlab= "Age (yrs)" , ylab= "Frequency", col="cadetblue", col.main="cadetblue", las=1)
```
Describe number 10:Now, with the same dataset and x and y variable, We can customize this Histogram by adding colours to the bars as well as to the Title with choosed "cadetblue" for both of them.

11. Create an overdense plot using the same specifications as the previous scatterplot. But,
include two levels of shading for the “black” data points. Title the plot “Plot 3: Harrison
Density Shaded by Era"

```{r}

  plot (data_f$fl, data_f$age, xlim=c(0, 500), ylim=c(0,15), xlab= "Fork Length (mm)",
      ylab= "Age(yrs)", main="Plot 3 : Harrison Density Shaded by Era",  col=c("black", "black")  , cex=1,pch=3, las=1 )


```
Describe number 11:This ScatterPlot shows the strong,positive and Linear relationship between Fork length and Age
of the Dataset: data1_Harrison, with including two levels of shading for the “black” data points.


12. Create a new object called “tmp” that includes the first 3 and last 3 records of the whole
data set

```{r}
tmp <- data.frame(rbind(head(data1,3),tail(data1, 3)))
tmp
```
Describe number 12:Here, Create a new Dataframe(tmp) with two dataframes (First 3 Records which should used head and Last 3 records (tail) of BullTroutRML2 dataset which is called (data1), having each values in the column.

13. Display the “era” column in the new “tmp” object
```{r}
eratmp <- tmp$era 
eratmp
```
Describe number 13:Create new object called(eratmp) to show the "era" column in the new "tmp" object.

14.Create a pchs vector with the argument values for + and x. Then create a cols vector
with the two elements “red” and “gray60”
```{r}
pch <-c("+", "x")
pch
cols <-c ("red", "gray60")
cols
```
Describe number 14: Now in this section, i have included the Shape of the Points and Colour of the Points in vectors - pchs and
cols respectively.

15. Convert the tmp object values to numeric values. Then create a numeric numEra object
from the tmp$era object
```{r}
tmp$era <- as.numeric(tmp$era)
tmp$era
 
numEra <- ("tmp$era") 
numEra

```
Describe number 15:Here, first we need to convert tmp object to numeric, so i created tmp$era and then use its function. Then as the question mentioned, i created a new numeric object called numEra from the tmp$era.

16. Associate the cols vector with the tmp era values
```{r}
cols <- tmp$era
cols
```

17. Create a plot of “Age (yrs)” (y variable) versus “Fork Length (mm)” (x variable) with the
following specifications:
• Limit of x axis is (0,500)
• Limit of y axis is (0,15)
• Title of graph is “Plot 4: Symbol & Color by Era”
• X axis label is “Age (yrs)”
• Y axis label is “Fork Length (mm)”
• Set pch equal to pchs era values
• Set col equal to cols era values

```{r}
plot(data1$fl,data1$age, main = "Plot 4: Symbol & Color by Era", xlim=c(0, 500), ylim=c(0,15), xlab="Fork Length", ylab = "Age(Yrs)", pchs=pch, cols=col)

```
Describe number 17:As per the above graph, This Scatter plot shows weak, positive and Linear relationship with no potential outliers between the Fork length and Age of the tmp Dataset.

18. Plot a regression line of the previous plot with a dashed line with width 2 and color
“cadetblue”
```{r}
plot(data1$fl,data1$age, main = "Plot 5: Regression Overlay", xlim=c(0, 500), ylim=c(0,15), xlab="Fork Length", ylab = "Age(Yrs)", pch = pchs, cols=col)

abline(lm(age~fl, data = data1), lty=2, lwd=2, col = 'cadetblue')
```
Describe number 18:Here, Only 1 Regression line has been overlayed.
The regression line which can be seen, makes it easy to understand the data, by showing how the data points approximately fit in the line perfectly.
This shows that Scatter plot here is showing Weak, Positive and Linear Relationship with no
Outliers between Fork length and Age of the tmp Dataset.


19. Place a legend of levels by era with pchs symbols in the top left of the plot with the
following specifications :
• Inset of 0.05
• No box around the legend
• Font size: 75% of nominal
```{r}
plot(data1$fl,data1$age, main = "Plot 6 : Legend Overlay", xlim=c(0, 500), ylim=c(0,15), xlab="Fork Length", ylab = "Age(Yrs)", pch = pchs, cols=col)
abline(lm(age~fl, data = data1), lty=1, lwd=2, col = 'red')
legend('topleft', inset = 0.05, title="Era",c("1977-80","1977-01"), bty='n', cex=0.75)

```
Describe number 19:Here, as we write topn left, the legend is appeared on that side; Legend Title - Era is displayed with
Color black : 1977-80 ( Shape Circle )
Color red : 1997- 01 (Shape circle )

Bibliography:

 https://r-coder.com/add-legend-r/
 
 geeksforgeeks.com
 
 Tutorialspoint.com
 
 
 Stackoverflow.com
 
 R in Action by R. Kabacoff
 
 Appendix:
 
>name <- ("Shamim Sherafati")

>r_name = paste("Plotting Basics : ",name)

>print(r_name)

>install.packages ("plyr")

>install.packages ("FSA")

>install.packages ("FSAdata")

>install.packages ("magrittr")

>install.packages ("dplyr")

>install.packages ("plotrix")

>install.packages ("ggplot2")

>install.packages ("moments")

>data1 <- read.table (file="~/desktop/BullTroutRML2.txt", header= TRUE, sep=",", >stringsAsFactors = FALSE)


>data1
>head (data1,  n=3 )

>tail (data1, n=3 )

>#headtail (data1, n=3)

>library(dplyr)

>data_f = filter(data1, lake == 'Harrison')

>data_f

>head(data_f, n=3)

>tail(data_f, n=3)

>str(data_f)

>t <- summary(data_f)

>t

>plot (data_f$fl, data_f$age, main = ("Plot 1: Horrison Lake Trout"), xlim = c(0, 500), ylim = c(0, >15), xlab = ("Fork Length (mm)"), ylab = ("Age (yrs)"),col="red", cex=0.5, pch=16)

>hist(data_f$age, main= "Plot 2: Harrison Fish Age Distribution" , xlab= "Age (yrs)" , ylab= >"Frequency", col="cadetblue", col.main="cadetblue", las=1)

>plot (data_f$fl, data_f$age, xlim=c(0, 500), ylim=c(0,15), xlab= "Fork Length (mm)",
      ylab= "Age(yrs)", main="Plot 3 : Harrison Density Shaded by Era",  col=c("black", "gray")  , cex=0.5,pch=15, las=1 )

>tmp <- data.frame(rbind(head(data1,3),tail(data1, 3)))

>tmp

>eratmp <- tmp$era 

>eratmp

>pch <-c("+", "x")

>pch

>cols <-c ("red", "gray60")

>cols

>tmp$era <- as.numeric(tmp$era)

>tmp$era
 
>numEra <- ("tmp$era") 

>numEra

>cols <- tmp$era

>cols

>plot(data1$fl,data1$age, main = "Plot 4: Symbol & Color by Era", xlim=c(0, 500), ylim=c(0,15), xlab="Fork Length", ylab = "Age(Yrs)", pch = pchs, cols=col)

>plot(data1$fl,data1$age, main = "Plot 5: Regression Overlay", xlim=c(0, 500), ylim=c(0,15), xlab="Fork Length", ylab = "Age(Yrs)", pch = pchs, cols=col)
abline(lm(age~fl, data = data1), lty=2, lwd=2, col = 'cadetblue')

>plot(data1$fl,data1$age, main = "Plot 6 : Legend Overlay", xlim=c(0, 500), ylim=c(0,15), xlab="Fork Length", ylab = "Age(Yrs)", pch = pchs, cols=col)
abline(lm(age~fl, data = data1), lty=1, lwd=2, col = 'red')
legend('topleft', inset = 0.05, title="Era",c("1977-80","1977-01"), bty='n', cex=0.75)


